Biases in Planet Occurrence Caused by Unresolved Binaries in Transit Surveys

### 2018-03-01 paper/main.tex
update title

\title{ Binarity's Biases on Transit Survey Occurrence Rates }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\affiliation{
    NASA Sagan Fellow}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars have binary companions. Unresolved and
unrecognized binaries give rise to systematic errors in planet
occurrence rates, including misclassified planets and mistakes in
completeness corrections.  The individual errors can have different
signs, making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive a
formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula depends
on the binary fraction; the mass function of the secondary stars; and
the true occurrence of planets around primaries, secondaries, and
single stars. It also takes into account the Malmquist bias by which
binaries are over-represented in flux-limited samples. Application of
the formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the errors are potentially larger: the occurrence of Earth-sized
planets could be overestimated by as much as 50\%. We also show that
whenever high-resolution imaging reveals a transit host star to be a
binary, the planet is usually more likely to orbit the primary star
than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

At the outset of this study it was not clear to us whether the neglect
of binaries is a serious problem, or even whether the net effect of
the errors is positive or negative.  The goals of this study were to
provide a framework for dealing with these issues, and to gauge at
least the order of magnitude of the systematic errors.  In this
spirit, our models are idealized.  We do not attempt a detailed
correction of the results from {\it Kepler} or any other real transit
survey.  We took inspiration from the insightful analytic model for
transit surveys by~\citet{pepper_using_2003}.

@@

From within this sample of apparently searchable sources, the
observers sum the number of detections ${\rm d}N_{\rm det}$ of planets
with an apparent radius between $r\a$ and $r\a + dr\a$.  There will be
three contributions to this sum,
\begin{align}
  {\rm d}N_{{\rm det},0} &= 
      \frac{N_\star(r\a) p_{\rm tra}}{1+\mu}
      \Gamma_0(r\a)\mathrm{d}r\a,
  \label{eq:n0} \\
  {\rm d}N_{{\rm det},1} &= 
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_1(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a),
	\label{eq:n1} \\
  {\rm d}N_{{\rm det},2} &=
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_2(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a).
	\label{eq:n2}
\end{align}
where here and elsewhere, the subscript 0 refers to single stars, 1
refers to primary stars of binaries, and 2 refers to secondary stars
of binaries. The apparent occurrence rate density is therefore
\begin{equation}
  \Gamma\a(r\a) = \frac{{\rm d}N_{\rm det}}{{\rm d}r\a} \frac{1}{N_\star(r\a) p_{\rm tra}} =
  \frac{\Gamma_0(r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_1(\sqrt{2}r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_2(\sqrt{2}r\a)}{1+\mu}.
	\label{eq:formula_twin_binary}
\end{equation}
Our earlier results, Equations~\ref{eq:rate_at_rp}
and~\ref{eq:correction_diluted_rp}, are recovered by integrating this
formula over $r$ after inserting
\begin{equation}
  \Gamma_i(r) = \Lambda_{r\p}\, \hat{\delta}(r-r\p),~~{\rm for}\ i\in\{0,1,2\}
\end{equation}
where $\hat{\delta}$ is the Dirac delta function, and $\Lambda_{r\p}$
is the number of planets per star with size $r\p$ considered in
Section~\ref{sec:model_1}.

@@

To compute the apparent occurrence rate density, we need to make
the following modifications to Equation~\ref{eq:formula_twin_binary}:
%
\begin{enumerate}
%
\item The Malmquist bias is different.  For a given value of $r\a$, the
  number of binary systems in the searchable sample with mass ratio
  $(q, q+\mathrm{d}q)$ is
\begin{equation}
    {N_\star(r\a)\over{1+\mu}}\, {n_{\rm b}\over n_{\rm s}}
    \left[{L(M) + L(qM) \over L(M)}\right]^{3/2}
    f(q)\,\mathrm{d}q,
    \label{eq:N_binaries}
\end{equation}
where $\mu$ is given by
\begin{equation}
    \mu= 
    \int_0^1 {n_{\rm b}\over n_{\rm s}}\left[{L(M) + L(qM)
    \over L(M\a)}\right]^{3/2} f(q)\,\mathrm{d}q.
    \label{eq:mu_general}
\end{equation}
Equation~\ref{eq:N_binaries} replaces ${\mu N_\star(r\a)/(1+\mu)}$
in Equations~\ref{eq:n1} and~\ref{eq:n2}.
The whole equation is then integrated to give the number of
detected planets.

@@

Taking these modifications into account, a general formula for the
apparent rate density is
\begin{align}
    \notag
    \Gamma\a(r\a) &= \frac{1}{1+\mu} \,
    \left\{ \Gamma_0(r\a)+ 
    \frac{n\b}{n\s}
    \left[ \int_0^1 \mathrm{d}q \,
           \mathcal{D}_1^3 f(q)\cdot
           \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a)\,
    \right.   
    \right. \\
    & \quad\quad\quad\quad\quad \left.\left.
    +\int_0^1 {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\cdot \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a)\cdot
         \frac{R(qM)}{R(M)} q^{-1/3}
    \right] \right\}.
    \label{eq:general_Gamma_a}
\end{align}
We note that the combination $\mathcal{D}_1^3f(q)$ is the mass-ratio
distribution for binaries contained within the searchable sample of
sources.  With this in mind we may rewrite
Equation~\ref{eq:general_Gamma_a} as
\begin{align}
    \Gamma\a(r\a)
    =
    {1\over 1+\mu} \,
    \left[
       \Gamma_0(r\a)
       +\mu\,
       \left\langle
       \mathcal{D}_1\cdot\Gamma_1\left(\mathcal{D}_1r\a\right)
       +
       {\mathcal{D}_2}\cdot\Gamma_2\left(\mathcal{D}_2r\a\right)
       \cdot
       {R(qM) \over R(M)}q^{-1/3}
       \right\rangle
    \right],
\end{align}
where the angle brackets denote averaging over all the binaries in the
searchable sample.

@@

The situation for smaller apparent radii is more nuanced.  For
$Z_2/Z_0=0$ or 0.5, any transit signals from binaries are still always
more likely to arise from the primary star.  However, if planets exist
at the same rate in primaries and secondaries ($Z_2/Z_0=1$), then
below apparent radii of $0.4\,R_\oplus$, more of the detected planets
in binaries come from secondaries.  They are actually much larger
planets for which $\delta$ has been substantially reduced by the light
from the primary star.  As we consider apparent radii ranging from 2
to 1.4\,$R_\oplus$, the fraction of detected planets with binary
companions increases by anywhere from $6\%$~to~$12\%$, depending on
the relative occurrence of planets about primaries and secondaries.

@@

\paragraph{Hot Jupiters} Another frequently discussed problem
involving occurrence rates is the discrepancy between the occurrence
rates of hot Jupiters that have been measured in different surveys, as
summarized in Table~\ref{tab:hj_rates}.  In particular, the California
Planet Search (CPS), a Doppler survey, found a higher rate ($12\pm 2$
per thousand stars) than the {\it Kepler} transit survey
($5.7^{+1.4}_{-1.2}$).  It has been suggested that unresolved binaries
in the {\it Kepler} sample might be an important factor, but the
calculations presented in Section~\ref{sec:further_models} suggest
that this is not the case.  We found that the effects of unresolved
binaries are likely to result in an overestimation of the hot-Jupiter
occurrence rate, which would worsen the discrepancy, rather than
relieve it. Some other possibilities are errors in stellar
classification of the {\it Kepler} stars, as well as differences in
the metallicity of the stars in each survey.  The occurrence rate of
hot Jupiters is known to vary strongly with stellar metallicity, with
$\Lambda \propto 10^{3.4{\rm [M/H]}}$ according to a recent
study~\citep{petigura_CKS_2017}.  \citet{guo_metallicity_2017} found
the mean metallicity of the {\it Kepler} stars to be [M/H]~$=-0.045\pm
0.009$, which is only slightly lower than that of the stars in the
Doppler survey, $-0.005\pm 0.006$.  Nevertheless, given the strong
dependence of occurrence on metallicity, even a difference of
$0.04$~dex could account for a factor of $10^{3.4\cdot 0.04} = 1.37$
in the occurrence rate.  After correcting for metallicity, the
discrepancy between the {\it Kepler} and CPS occurrence rates is
reduced to the 1.7$\sigma$ level.

>>>>

\title{ Biases in Planet Occurrence Caused by Unresolved Binaries in
        Transit Surveys }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\affiliation{
    NASA Sagan Fellow}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars have binary companions. Unresolved and
unrecognized binaries give rise to systematic errors in planet
occurrence rates, including misclassified planets and mistakes in
completeness corrections.  The individual errors can have different
signs, making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive a
formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula depends
on the binary fraction; the mass function of the secondary stars; and
the true occurrence of planets around primaries, secondaries, and
single stars. It also takes into account the Malmquist bias by which
binaries are over-represented in flux-limited samples. Application of
the formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the errors are potentially larger: the occurrence of Earth-sized
planets could be overestimated by as much as 50\%. We also show that
whenever high-resolution imaging reveals a transit host star to be a
binary, the planet is usually more likely to orbit the primary star
than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

At the outset of this study it was not clear to us whether the neglect
of binaries is a serious problem, or even whether the net effect of
the errors is positive or negative.  The goals of this study were to
provide a framework for dealing with these issues, and to gauge at
least the order of magnitude of the systematic effects.  In this
spirit, our models are idealized.  We do not attempt a detailed
correction of the results from {\it Kepler} or any other real transit
survey.  We took inspiration from the insightful analytic model for
transit surveys by~\citet{pepper_using_2003}.

@@

From within this sample of apparently searchable sources, the
observers sum the number of detections ${\rm d}N_{\rm det}$ of planets
with an apparent radius between $r\a$ and $r\a + dr\a$.  There will be
three contributions to this sum,
\begin{align}
  {\rm d}N_{{\rm det},0} &= 
      \frac{N_\star(r\a) p_{\rm tra}}{1+\mu}
      \Gamma_0(r\a)\mathrm{d}r\a,
  \label{eq:n0} \\
  {\rm d}N_{{\rm det},1} &= 
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_1(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a),
	\label{eq:n1} \\
  {\rm d}N_{{\rm det},2} &=
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_2(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a).
	\label{eq:n2}
\end{align}
where here and elsewhere, the subscript 0 refers to single stars, 1
refers to primary stars of binaries, and 2 refers to secondary stars
of binaries. The apparent occurrence rate density is therefore
\begin{equation}
  \Gamma\a(r\a) = \frac{{\rm d}N_{\rm det}}{{\rm d}r\a} \frac{1}{N_\star(r\a) p_{\rm tra}} =
  \frac{\Gamma_0(r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_1(\sqrt{2}r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_2(\sqrt{2}r\a)}{1+\mu}.
	\label{eq:formula_twin_binary}
\end{equation}
Our earlier results, Equations~\ref{eq:rate_at_rp}
and~\ref{eq:correction_diluted_rp}, are recovered by integrating this
formula over $r$ after letting the respective rate densities be
\begin{equation}
  \Gamma_i(r) = \Lambda_{r\p}\, \hat{\delta}(r-r\p),~~{\rm for}\ i\in\{0,1,2\}
\end{equation}
where $\hat{\delta}$ is the Dirac delta function, and $\Lambda_{r\p}$
is the number of planets per star with size $r\p$ considered in
Section~\ref{sec:model_1}.

@@

To compute the apparent occurrence rate density, we need to make
the following modifications to Equation~\ref{eq:formula_twin_binary}:
%
\begin{enumerate}
%
\item The Malmquist bias is different.  For a given value of $r\a$, the
  number of binary systems in the searchable sample with mass ratio
  $(q, q+\mathrm{d}q)$ is
\begin{equation}
    {N_\star(r\a)\over{1+\mu}}\, {n_{\rm b}\over n_{\rm s}}
    \left[{L(M) + L(qM) \over L(M)}\right]^{3/2}
    f(q)\,\mathrm{d}q,
    \label{eq:N_binaries}
\end{equation}
where $\mu$ is given by
\begin{equation}
    \mu= 
    \int_0^1 {n_{\rm b}\over n_{\rm s}}\left[{L(M) + L(qM)
    \over L(M)}\right]^{3/2} f(q)\,\mathrm{d}q.
    \label{eq:mu_general}
\end{equation}
Equation~\ref{eq:N_binaries} replaces ${\mu N_\star(r\a)/(1+\mu)}$
in Equations~\ref{eq:n1} and~\ref{eq:n2}.
The whole equation is then integrated to give the number of
detected planets.

@@

Taking these modifications into account, a general formula for the
apparent rate density is
\begin{align}
    \notag
    \Gamma\a(r\a) &= \frac{1}{1+\mu} \,
    \left\{ \Gamma_0(r\a)+ 
    \frac{n\b}{n\s}
    \left[ \int_0^1 \mathrm{d}q \,
           \mathcal{D}_1^3 f(q)\cdot
           \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a)\,
    \right.   
    \right. \\
    & \quad\quad\quad\quad\quad \left.\left.
    +\int_0^1 {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\cdot \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a)\cdot
         \frac{R(qM)}{R(M)} q^{-1/3}
    \right] \right\}.
    \label{eq:general_Gamma_a}
\end{align}
We note that the combination $\mathcal{D}_1^3f(q)$ is the mass-ratio
distribution for binaries contained within the searchable sample of
sources.  With this in mind we may rewrite
Equation~\ref{eq:general_Gamma_a} as
\begin{align}
    \Gamma\a(r\a)
    =
    {1\over 1+\mu} \,
    \left[
       \Gamma_0(r\a)
       +\mu\,
       \left\langle
       \mathcal{D}_1 \, \Gamma_1\left(\mathcal{D}_1r\a\right)
       +
       {\mathcal{D}_2} \, \Gamma_2\left(\mathcal{D}_2r\a\right)
       \,
       {R(qM) \over R(M)}q^{-1/3}
       \right\rangle
    \right],
\end{align}
where the angle brackets denote averaging over all the binaries in the
searchable sample.

@@

The situation for smaller apparent radii is more nuanced.  For
$Z_2/Z_0=0$ or 0.5, any transit signals from binaries are still always
more likely to arise from the primary star.  However, if planets exist
at the same rate in primaries and secondaries ($Z_2/Z_0=1$), then
below apparent radii of $0.4\,R_\oplus$, more of the detected planets
in binaries come from secondaries.  They are actually much larger
planets for which the transit signal amplitude has been substantially
reduced by the light from the primary star.  As we consider apparent
radii ranging from 2 to 1.4\,$R_\oplus$, the fraction of detected
planets with binary companions increases by anywhere from
$6\%$~to~$12\%$, depending on the relative occurrence of planets about
primaries and secondaries.

@@

\paragraph{Hot Jupiters} Another frequently discussed problem
involving occurrence rates is the discrepancy between the occurrence
rates of hot Jupiters that have been measured in different surveys, as
summarized in Table~\ref{tab:hj_rates}.  In particular, the California
Planet Search (CPS), a Doppler survey, found a higher rate ($12\pm 2$
per thousand stars) than the {\it Kepler} transit survey
($5.7^{+1.4}_{-1.2}$).  It has been suggested that unresolved binaries
in the {\it Kepler} sample might be an important factor, but the
calculations presented in Section~\ref{sec:further_models} suggest
that this is not the case.  We found that the effects of unresolved
binaries are likely to result in an overestimation of the hot-Jupiter
occurrence rate, which would worsen the discrepancy, rather than
relieve it. Some other possibilities are errors in stellar
classification of the {\it Kepler} stars, as well as differences in
the metallicity of the stars in each survey.  The occurrence rate of
hot Jupiters is known to vary strongly with stellar metallicity, with
$\Lambda \propto 10^{3.4{\rm [M/H]}}$ according to a recent
study~\citep{petigura_CKS_2017}.  \citet{guo_metallicity_2017} found
the mean metallicity of the {\it Kepler} stars to be [M/H]~$=-0.045\pm
0.009$, which is only slightly lower than that of the stars in the
Doppler survey, $-0.005\pm 0.006$.  Nevertheless, given the strong
dependence of occurrence on metallicity, even a difference of
$0.04$~dex could account for a factor of $10^{3.4(0.04)} = 1.37$
in the occurrence rate.  After correcting for metallicity, the
discrepancy between the {\it Kepler} and CPS occurrence rates is
reduced to the 1.7$\sigma$ level.
